---
title: "a_b_test"
author: "Andrew Cotter"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# A/B Testing Analysis Using an Example Data Set

## Introduction

### Scenario

A fast-food chain plans to add a new item to its menu. However, they are still undecided between three possible marketing campaigns for promoting the new product. In order to determine which promotion has the greatest effect on sales, the new item is introduced at locations in several randomly selected markets. A different promotion is used at each location, and the weekly sales of the new item are recorded for the first four weeks.

### Goal

Evaluate A/B testing results and decide which marketing strategy works the best.

### Columns

-   **MarketID**: unique identifier for market
-   **MarketSize**: size of market area by sales
-   **LocationID**: unique identifier for store location
-   **AgeOfStore**: age of store in years Promotion: one of three promotions that were tested
-   **week**: one of four weeks when the promotions were run
-   **SalesInThousands**: sales amount for a specific LocationID, Promotion, and week

Data was sourced from [Kaggle](https://www.kaggle.com/datasets/chebotinaa/fast-food-marketing-campaign-ab-test)

------------------------------------------------------------------------

## Data Loading and Inspection

```{r echo=F}
library(readxl)
d = read.csv("WA_Marketing-Campaign.csv")
head(d)
```

Upon inspection, I notice that there are a few columns that need redefining in terms of their data type. MarketID, LocationID, and Promotion should all be factors rather than integers.

```{r}
d$MarketID = as.factor(d$MarketID)
d$LocationID = as.factor(d$LocationID)
d$Promotion = as.factor(d$Promotion)

print(summary(d))
print(range(table(d$MarketID)))
```

It appears that there are some imbalances in the data set.

-   For MarketID, there are a variety of different counts that range from 24-88. The number of observations that we have per market is going to influence the degree of confidence in the conclusions that we can draw within that market.
-   In terms of market sizes, it appears that medium markets are by far the most common. Again, we are going to have less confidence in the conclusions that we draw about large and especially small markets compared to the ones we draw about medium markets.
-   Finally, it appears that promotion 1 was shown slightly less than promotions 2 and 3.

```{r}
table(is.na(d))
```

There are no missing values.

------------------------------------------------------------------------

## Searching for Confounds

Before we go about building a model for our variable of interest (Promotion), we need to do some checks to ensure that there are no *confounds* that were introduced by the experimental design.

A *confound* is any variable, whether it is in the data set or not, that impacts both the treatment and the outcome. This commonly arises from poor experimental design, such as imbalances in the characteristics of the control and treatment groups.

### Market Size

First, we will investigate market size. To reiterate, we need to establish that market size is a good predictor of both the outcome and the treatment in order to conclude that it is a confound. Starting with its impact on the outcome:

```{r, echo = F}
library(dplyr)

d %>%
  group_by(MarketSize) %>%
  summarise(avg_sales = mean(SalesInThousands))

(summary(aov(SalesInThousands~MarketSize, d)))
```

The data is suggesting that the market size has an appreciable impact on sales. The 1-way ANOVA suggests statistical significance, and the summary table shows that large markets see, on average, about 60% more in sales when compared to medium markets.

Next, we need to establish that promotion campaigns are unequally distributed across the different market sizes in order to deem market size a confound. Since we have such differences in the number of observations per market, I will use a proportion table to produce the point estimates, and I will use the chi-square test to check for equality across the different cells of the table (of counts).

```{r, echo = F}
#Proportion table for percentage of market sizes per promotion type
round(
  prop.table(
    table(d$Promotion, d$MarketSize),
    margin = 1),
  2
)

#Chi-Square test for equality of observations per cell
chisq.test(table(d$Promotion, d$MarketSize))
  
```
Based on this analysis, I will not worry too much about market size being a confound. The distribution of promotions is balanced enough across the different markets, based on the insignificance of the chi-squre value.

Next, in our further exploration of possible confounds, we can check to see if the age of the store has an impact on sales.

```{r, echo = F}
library(ggplot2)
ggplot(d, aes(x = AgeOfStore, y = SalesInThousands))+
  geom_jitter(width = 0.4, alpha = 0.2, color = "blue", size = 3)+
  theme_classic()
cor.test(d$AgeOfStore, d$SalesInThousands)
```
